---
title: "Sample_Maps"
author: "Vaibhav Agarwal"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Making Maps

Here is a Rmnd file you can use to make maps! 

* sf is the main package we use to make maps
* sf accepts data in many formats -- mainly shapefiles
* shape files are downlaoded as zipfiles and you must open the entire zip file for it to work :) 
* For this example, only .shp for finland is enough, in the future it would be required to use zipfiles or it is also possible to save all the files in the same folder



```{r}
# Sf is the main file 
library(sf)
library(ggplot2)
library(dplyr)
library(ggthemes)


# this is to set the path of your shapefile (must have all other files in the same folder as well)
fin_mun <- read_sf("C:/Users/User/OneDrive - University of Helsinki/Desktop/FIN_adm4.shp")
# plot(fin_mun) # You can try this as well to see all the plots. use max.plot=N  to check out all the columns


# Incase you have smaller regions and want group them up into larger categories. Better categories exist in the data, this is just for trying.

category_lookup <- list(
  `Capital Region` = c("Eastern Uusimaa"),
  `Uusimaa` = c("Uusimaa"),
  `Southern Finland` = c("Kymenlaakso", "Päijänne Tavastia", "Tavastia Proper", "South Karelia"),
  `Western Finland` = c("Finland Proper", "Satakunta", "Pirkanmaa", "Southern Ostrobothnia","Ostrobothnia", "Central Ostrobothnia"),
  `Northern and Eastern Finland` = c("Central Finland", "North Karelia", "North Savonia","Northern Ostrobothnia", "Kainuu", "Lapland", "Southern Savonia"),
  `Åland Islands` = c()
)

# Using the stack fucntion to attach the list to make the dictionary into a sort of df
# How stack works:
## example we have
### column names a,b 
### row1 - 1,2
### row2 -4,5
### row3 - 9,3

# stacking will make this 
## row1 - 1,a 
## row2 - 4,a 
## row3 - 9,a 
## row4 - 2,b
## row5 - 5,b
## row6 -3,b 

#it is the sane as converting from wide to long.

category_df <- stack(category_lookup)
colnames(category_df) <- c("Region", "Category") #renaming 

# lets add the columnn. 
# we will do so by merging 
# the common column is Name_2 and Region (the rows have same info)
# We merge them together and want only those which merge to be in df, si using all.x = TRUE
fin_mun <- merge(fin_mun, category_df, by.x = "NAME_2", by.y = "Region", all.x = TRUE)


#Quick Plotting
ggplot() + 
   geom_sf(aes(fill = fin_mun$Category),color = 'black',data = fin_mun)

```
## Some quick map making help

### Shading only some subsets 

```{r}
#If you want for some instance only to shade southern finland and capital region and leave the rest blank, 
# chose the color of the empty df. by default it is grey. If you dont want that pass this line of code 

ggplot() +
  geom_sf(fill = 'gold',color = 'black',data = fin_mun)+ 
  theme_void()

# lets say you want to shade capital region and southern finland 
# step 1 - make subsets
cap_region <- fin_mun %>% 
  filter(Category == 'Capital Region')

south_FI <- fin_mun %>% 
  filter(Category == 'Southern Finland')

# After this, overlay these graphs one over the other
#Next, coloring always have sequencially. If you pass the entire df at the end, the whole map will be yellow, so pass them in order of detail. 
# like this
ggplot() +
  geom_sf(fill = 'gold',color = 'black',data = fin_mun)+
  geom_sf(fill = 'red',color = 'black',data = south_FI)+
  geom_sf(fill = 'black',color = 'black',data = cap_region)+
  theme_void()

# You can also add some title, fommatting and labels to your graph to make it more beautiful!

ggplot() +
  geom_sf(fill = 'gold', color = 'black', data = fin_mun) +
  geom_sf(fill = 'red', color = 'black', data = south_FI) +
  geom_sf(fill = 'black', color = 'black', data = cap_region)+
  theme_void() +
  labs(title = 'Adding colors and shading subsets on maps in R',
       subtitle = 'Entire Finland, Southern Finland, Capital Region') +
  theme(
    plot.title = element_text(color = "black", size = 12, hjust = 0.5),
    plot.subtitle = element_text(color = 'black', size = 8, hjust = 0.5),
    plot.caption = element_text(face = "italic"),
    legend.position = 'none'
  )



```

## Merging this data with a random column and plotting these graphs on a color scale 

```{r}
# Megre the column based on common place name. 
# For example, of NAME_2 and your df both have 'uusimaa' exactly as it is, then you can merge the 2 adding freq in the other geo dataframe
#If they are not same 'uusimaa' vs 'Uusimaa', then they will not merge. 
# It is best to first convert to lowercase and check for similarity 
# Then you can do left merge, i.e., everything in df1 will be considered important and those rows wont merge
#I will look something like this 

### 
#fin_mun <- merge(fin_mun, 'your df', by.x = "your column in geo df", by.y = "freq column", all.x = TRUE)
###

#For now, I will directly create
fin_mun$value <- as.integer(runif(nrow(fin_mun), 0, 10))

ggplot() +  
   geom_sf(aes(fill = value),color = 'black',linewidth = 0.6,data = fin_mun)+
  theme_void()+
  theme(legend.position = 'right') +
  scale_fill_viridis_c(option = "inferno") +
  labs(title = 'Adding color based on frequency', 
    subtitle = '',
    fill = 'Random Variable')+
  theme(
    plot.title = element_text(color="black", size=12, hjust = 0.5),
    plot.subtitle = element_text(color = 'black',size = 8, hjust = 0.5),
    plot.caption = element_text(face="italic"),
    legend.position = 'right')
    
```


## Combining 1 and 2. Making a zoomed in map with frequency  

```{r}

west_FI <- fin_mun %>% 
  filter(Category == 'Western Finland')

ggplot() +  
  #geom_sf(fill = 'white',color = 'black', linewidth = 0.6,data = fin_mun)+ #adding defualt color to rest. Remove # for full map
   geom_sf(aes(fill = value),color = 'black',linewidth = 0.6,data = west_FI)+ #using south FO file we made sometime back 
  theme_void()+
  scale_fill_viridis_b(option = 'turbo')+
  theme(legend.position = 'right') +
  labs(title = 'Adding color based on frequency', 
    subtitle = 'This also includes subsets',
    fill = 'Random Variable')+
  theme(
    plot.title = element_text(color="black", size=12, hjust = 0.5),
    plot.subtitle = element_text(color = 'black',size = 8, hjust = 0.5),
    plot.caption = element_text(face="italic"),
    legend.position = 'right')

```
